---
title: "gi-ws-09-1_forest_stuff"
author: "Charlotte Gohr, Marvin Ludwig, Klaas Rueggeberg"
date: "31 Januar 2017"
output: html_document
---

## Part 1: Crop input to same area

```{r, eval = FALSE}
# gi-ws-09-1 forest related stuff
# preprocessing rasters: crop all to the same extent

# initialise script
source("d:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/init.R")
path <-  fun_init("gi", "09")

library(raster)

# load input rasters
input <- lapply(list.files(path$gi$input, full.names =  TRUE), function(x){
  stack(x)
})
names(input) <- list.files(path$gi$input)

# crop every raster to the extend of the aerial
crops <- lapply(seq(length(input)), function(i){
  if(input[[i]]@extent != input$geonode_ortho_muf_1m.tif@extent){
    return(crop(input[[i]], input$geonode_ortho_muf_1m.tif))
  }else{
    return(input[[i]])
  }
})
names(crops) <- names(input)

saveRDS(crops, file = paste0(path$gi$RData, "muf.rds"))

```


## Part 2: Identify trees
```{r, eval = FALSE}
library(link2GI)
library(raster)
library(rgrass7)
library(gdalUtils)
library(rgdal)

# load preprocessed rasters
input <- readRDS(paste0(path$gi$RData, "muf.rds"))

# init GRASS
linkGRASS7(x = input$lidar_dem_01m.tif, searchPath = "C:\\GIS\\QGIS")

# canopy height model
# trees are "hanging" upside down, the smaller the number, the larger the tree!
chm <- input$lidar_dsm_01m.tif - input$lidar_dem_01m.tif
chm[chm < 0] <- 0
chm <- (chm*(-1)) + chm@data@max
# remove structures smaller than 3m
chm[chm > maxValue(chm)-3] <- maxValue(chm)
writeRaster(chm, filename = paste0(path$gi$run, "chm.tif"), overwrite = TRUE)

# SAGA Channel Network and Drainage Basins
gdalwarp(srcfile = paste0(path$gi$run, "chm.tif"),
         dstfile = paste0(path$gi$run, "chm.sdat"),
         of = "SAGA", overwrite = TRUE)

system(paste0("saga_cmd ta_channels 5",
              " -DEM=", path$gi$run,"chm.sgrd",
              " -CONNECTION=", path$gi$run,"trees.sgrd",
              " -THRESHOLD=1"))

gdalwarp(srcfile = paste0(path$gi$run, "trees.sdat"),
         dstfile = paste0(path$gi$run, "potential_trees.tif"),
         of = "GTiff", overwrite = TRUE)


# # # clean up some false trees # # #
pot_trees <- raster(paste0(path$gi$run, "potential_trees.tif"))

# every pixel with a value larger than 4 is a tree
pot_trees[pot_trees < 6] <- NA

# smooth lcc and compare classes with trees
# forest ids are: 2,3,10,11,12,20,21,22,30,40,41,50,51,52,72
smooth_lcc <- focal(unstack(input$geonode_muf_lcc_prediction.tif)[[1]], matrix(1,3,3), modal)

id_forest <- c(2,3,10,11,12,20,21,22,30,40,41,50,51,52,72)
pot_trees[!(smooth_lcc %in% id_forest)] <- NA
writeRaster(pot_trees, filename = paste0(path$gi$run, "trees.tif"), overwrite = TRUE)

# GRASS: create point vector
execGRASS("r.import", flags = c("o","overwrite") , 
          parameters = list(input = paste0(path$gi$run, "trees.tif"),
                            output = "trees"))

execGRASS("r.to.vect", flags = c("overwrite"),
          parameters = list(input = "trees", output = "point_trees", type = "point", column = "nodes"))

execGRASS('v.out.ogr',  flags = c("overwrite"),
          parameters = list(input = "point_trees",
                            output = paste0(path$gi$run, "trees.shp"),
                            format = "ESRI_Shapefile"))

# # # now the crowns # # #

# SAGA Watershed Segmentation
system(paste0("saga_cmd imagery_segmentation 0",
              " -GRID=", path$gi$run, "chm.sgrd",
              " -SEGMENTS=", path$gi$run, "trees_segmentation.sgrd",
              " -BORDERS=NULL",
              " -OUTPUT=0",
              " -DOWN=0",
              " -JOIN=0",
              " -THRESHOLD=0.000000",
              " -EDGE=1",
              " -BBORDERS=0"))
# convert to vector
system(paste0("saga_cmd shapes_grid 6",
              " -GRID=", path$gi$run, "trees_segmentation.sgrd",
              " -POLYGONS=", path$gi$run, "tree_crowns.shp",
              " -CLASS_ALL=1",
              " -SPLIT=0",
              " -ALLVERTICES=0"))

# now we only want crowns that match our trees from earlier
# GRASS v.select
execGRASS("v.import", flags = "overwrite",
          parameters = list(input = paste0(path$gi$run, "tree_crowns.shp"), output = "crowns"))
execGRASS("v.import", flags = "overwrite",
          parameters = list(input = paste0(path$gi$run, "trees.shp"), output = "point_trees"))
execGRASS("v.select", flags = "overwrite",
          parameters = list(ainput = "crowns", binput = "point_trees", output = "real_crowns",
                            operator = "contains"))
execGRASS("v.out.ogr", flags = "overwrite",
          parameters = list(input = "real_crowns",
                            output = paste0(path$gi$run, "real_crowns.shp"),
                            format = "ESRI_Shapefile"))


```


## Part 3: Fill trees with information
```{r, eval = FALSE}
# load preprocessed rasters
input <- readRDS(paste0(path$gi$RData, "muf.rds"))

# load trees and crowns from part 2
trees <- readOGR(paste0(path$gi$input, "trees.shp"))
crowns <- readOGR(paste0(path$gi$input, "crowns.shp"))

# init GRASS
linkGRASS7(x = input$lidar_dem_01m.tif, searchPath = "C:\\GIS\\QGIS")

# clean up attribute df
names(trees@data) <- c("id_tree", "nodes")
trees@data$id_crown <- NA

crowns@data$cat <- NULL
crowns@data$NAME <- NULL
crowns@data$ID <- seq(length(crowns))
names(crowns@data) <- c("id_crown", "id_tree")
crowns@data$id_tree <- NA
crowns@data$area <- NA

writeOGR(trees, dsn = paste0(path$gi$run, "trees.shp"), driver = "ESRI Shapefile", layer = "trees" ,overwrite = TRUE)
writeOGR(crowns, dsn = paste0(path$gi$run, "crowns.shp"), driver = "ESRI Shapefile", layer = "crowns", overwrite = TRUE)

# # # tree height and type # # #
# tree height
chm <- input$lidar_dsm_01m.tif - input$lidar_dem_01m.tif
writeRaster(chm, filename = paste0(path$gi$run, "tree_height.tif"), overwrite = TRUE)

# import all needed maps in GRASS
execGRASS("r.import", flags = c("o", "overwrite"),
          parameters = list(input = paste0(path$gi$run, "tree_height.tif"),
                            output = "tree_height"))
execGRASS("r.import", flags = c("o", "overwrite"),
          parameters = list(input = paste0(path$gi$run, "smooth_lcc3.tif"),
                            output = "lcc"))
execGRASS("v.import", flags = c("o", "overwrite"),
          parameters = list(input = paste0(path$gi$run, "trees.shp"),
                            output = "trees"))
execGRASS("v.import", flags = c("o", "overwrite"),
          parameters = list(input = paste0(path$gi$run, "crowns.shp"),
                            output = "crowns"))


# # # querying: get values from raster at points # # #
execGRASS("r.what", flags = c("overwrite", "n"),
          parameters = list(map = c("tree_height","lcc"),
                            points = "trees",
                            output = paste0(path$gi$run, "heights.csv"),
                            separator = "comma"))

heights <- read.csv(paste0(path$gi$run, "heights.csv"), header = TRUE, sep = ",", dec = ".")
# legend for lcc
type <- read.csv(paste0(path$gi$run, "tree_type.csv"), header = TRUE, sep = ",")

# attach the right type, age and height to attribute table
heights$type <- type$type[match(heights$lcc, type$id)]
heights$age <- type$age[match(heights$lcc, type$id)]

trees@data$height <- heights$tree_height
trees@data$tree_type <- heights$type
trees@data$age <- heights$age

head(trees@data)

# # # match crowns with trees # # #
# testing second query possibility which directly updates the column
execGRASS("v.what.vect", parameters = list(map = "trees",
                                           column = "id_crown",
                                           query_map = "crowns",
                                           query_column = "id_crown"))
# only the GRASS vector is updated; we need to update our SPDF
temp_trees <- readVECT(vname = "trees")
trees@data$id_crown <- temp_trees@data$id_crown
crowns@data$id_tree <- trees@data$id_tree[match(crowns@data$id_crown, trees@data$id_crown)]

# save both shapefiles
writeOGR(trees, dsn = paste0(path$gi$output, "trees.shp"), driver = "ESRI Shapefile", layer = "trees" ,overwrite = TRUE)
writeOGR(crowns, dsn = paste0(path$gi$output, "crowns.shp"), driver = "ESRI Shapefile", layer = "crowns", overwrite = TRUE)


# tree density heatmap
execGRASS("v.kernel", parameters = list(input = "trees", output = "tree_dens", radius = 20, kernel = "gaussian"))
execGRASS("r.out.gdal", flags = "overwrite",
          parameters = list(input = "tree_dens",
                            output = paste0(path$gi$output, "tree_denisty.tif"),
                            format = "GTiff"))

```

## Part 4: Show results
```{r, message=FALSE, warning=FALSE}
source("d:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/init.R")
path <-  fun_init("gi", "09")

library(rgdal)
library(raster)

trees <- readOGR(paste0(path$gi$output, "trees.shp"))
crowns <- readOGR(paste0(path$gi$output, "crowns.shp"))

# a look at the attribute tables
head(trees@data)


# density heatmap
dens <- raster(paste0(path$gi$output, "tree_denisty.tif"))
plot(dens)


```

#### Idee zur Biomasse:
Baumhöhe und Kronenfläche sind pro Baum vorhanden. Damit kann man analog zur Formel aus He et al. 2013 (siehe letztwöchige Aufgabe) die Biomasse pro Baum berechnen. Um zu spezifizieren könnte man die zugewiesene Baumart und Alter in die Berechnung mit einfließen lassen.





