---
title: "gi-ws-07-1"
author: "Charlotte Gohr, Marvin Ludwig, Klaas Rueggeberg"
date: "19 Dezember 2016"
output: html_document
---
## task / script overview

DEM Simulation: 

  + Preprocess DEM (important: fill sinks Wang & Liu)
  + Correct gauge based on maximum drainage in a buffered area
  + Calculate catchment area of that gauge
  + Crop the filled DEM to the catchment area
  + Run the Overland Flow Kinematic D8 module for several timesteps

How much rainfall?  

  + Create a raster with two cells and different values (so water can flow)
  + Create gauges at both cells
  + Observ what happens when running Overland Flow Kinematic D8 module

## script
```{r, eval = FALSE}
# gi-ws-07-1 waterflow#
# # # # # # # # # # # #

# initialize script
source("d:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/init.R")
path <- fun_init("gi", "07")

library(raster)
library(rgdal)
library(gdalUtils)
source(path$fun$fun_hy_analysis.R)
source(path$fun$overland_flow.R)

# # # # # # # # #

# new folders for the basic analysis and the flow simulation
path$hydro <- paste0(path$gi$run, "hydrology/")
path$two <- paste0(path$gi$run, "two_cells/")
path$sim <- paste0(path$gi$run, "overland_flow/")

# maybe create folders
#dir.create(path$hydro)
#dir.create(path$two)
#dir.create(path$sim)

# load DEM and get the filepath
DEM <- raster::raster(paste0(path$gi$input, "geonode_las_dtm_01m.tif"))
path$DEM <- DEM@file@name

# preprocess DEM: fill sinks (Wang & Liu)
hy_pre <- hy_basic_analysis(DEM_path = path$DEM, workdir = path$hydro)
print(summary(hy_pre))

# # # # # # # # # # # # # # # # # # # # 

# # # # Calculate catchment area for a given point (gauge = Pegel) # # # #
# # Correct position for gauge to biggest drainage (=Abfluss) in a buffered area # #

# create spatial point at estimate gauge
gauge <- data.frame(y = 50.840860, x = 8.684456, name = "gauge")
coordinates(gauge) <- ~ x + y

# WGS84_longlat because of the coordinate format; then transform to format of DEM
projection(gauge) <- CRS("+init=epsg:4326")
gauge <- spTransform(gauge, CRS("+init=epsg:25832"))

# extract values from the catchment_area raster around the estimate gauge
gauge_buffer <- as.data.frame(extract(hy_pre$catchment_area, gauge, buffer = 25, cellnumbers = TRUE)[[1]])

# position with biggest value is the new position of the gauge
gauge@coords <- xyFromCell(DEM, gauge_buffer$cell[which.max(gauge_buffer$value)])
# save as shapefile for input of the simulation
writeOGR(gauge, dsn = paste0(path$sim, "gauge.shp"), driver = "ESRI Shapefile", layer = "gauge", overwrite_layer = TRUE)


# # # # # # # # # # # # # # #

# # # # crop DEM to catchment area of corrected gauge # # # #
# Saga Upslope Area
upslope_output <- paste0(path$sim, "catchment_gauge.sdat")
system(paste0("saga_cmd ta_hydrology 4 ",
              " -TARGET_PT_X ", gauge@coords[1,1],
              " -TARGET_PT_Y ", gauge@coords[1,2],
              " -ELEVATION DEM_no_sinks.sdat",
              " -AREA ", upslope_output,
              " -METHOD 0", 
              " -CONVERGE=1.100000"))

# convert to tif
gdalwarp(srcfile = paste0(path$sim, "catchment_gauge.sdat"),
        dstfile = paste0(path$sim, "catchment_gauge.tif"),
        overwrite = TRUE,
        of = "GTiff")

# load catchment of the gauge tif
catchment_gauge <- raster(paste0(path$sim, "catchment_gauge.tif"))

# get nothing but the catchment area
catchment_gauge[catchment_gauge < 100] <- NA
catchment_gauge <- trim(catchment_gauge)
# crop the DEM
DEM_gauge <- crop(hy_pre$DEM_no_sinks, catchment_gauge)

# save croped DEM as sgrd and tif
writeRaster(DEM_gauge, filename = paste0(path$sim, "DEM_gauge.tif"), overwrite = TRUE)
writeRaster(DEM_gauge, filename = paste0(path$sim, "DEM_gauge.sdat"), overwrite = TRUE)

# # # # # # # # # # # # # # # # # # # # #

# # # # Create raster with two cells to get the percipitation values of the used module # # # #
# create raster with two cells
two <- raster(nrows = 1, ncols = 2, xmn = 0, xmx = 2, ymn=0, ymx = 1)
values(two) <- c(2000, 1)

# create gauges at both cells to observate the simulation
two_gauges <- as.data.frame(rasterToPoints(two))
coordinates(two_gauges) <- ~ x + y

# save both
writeRaster(two, filename = paste0(path$two, "two_cell.sdat"), overwrite  = TRUE)
writeOGR(two_gauges, dsn = paste0(path$two, "two_gauge.shp"), driver = "ESRI Shapefile", layer = "gauge", overwrite_layer = TRUE)


# # # # # # # # # # # # # # # # # #
# finally all input data is ready #
# # # # # # # # # # # # # # # # # #

# # # # run precipitation simulation # # # #

# two cells to test precipitation
sim_two_cells <- hy_overland_flow(DEM_path = paste0(path$two, "two_cell.sgrd"),
                                  gauges_path = paste0(path$two, "two_gauge.shp"),
                                  workdir = path$two,
                                  time = "1",
                                  steps = "1",
                                  roughness = "0")

# simulation with catchment area for three different time steps
timesteps <- c("0.1", "0.5", "1")

sim_catchment <- lapply(timesteps, function(x){
  output_dir <- paste0(path$sim, x, "/")
  
  hy_overland_flow(DEM_path = paste0(path$sim, "DEM_gauge.sgrd"),
                   gauges_path = paste0(path$sim, "gauge.shp"),
                   workdir = output_dir,
                   time = "24",
                   steps = x)
  
})
names(sim_catchment) <- timesteps

# save results
saveRDS(sim_two_cells, file = paste0(path$gi$run, "sim_two_cells.rds"))
saveRDS(sim_catchment, file = paste0(path$gi$run, "sim_catchment.rds"))
```


## Show results
```{r, include = FALSE}
source("d:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/init.R")
path <- fun_init("gi", "07")
path$hydro <- paste0(path$gi$run, "hydrology/")
path$two <- paste0(path$gi$run, "two_cells/")
path$sim <- paste0(path$gi$run, "overland_flow/")
library(raster)
```

### 1. How much rain fall?
```{r}
sim_two_cells <- readRDS(paste0(path$gi$run, "sim_two_cells.rds"))
print(sim_two_cells$gauges_flow)
```
A two cell simulation with one timestep gets us a total of 100 rain units, so every cell gets 50 rain units  

### 2. Different time steps
```{r}
sim_catchment <- readRDS(paste0(path$gi$run, "sim_catchment.rds"))

plot(sim_catchment$`0.1`$overland_flow)
plot(sim_catchment$`0.5`$overland_flow)
plot(sim_catchment$`1`$overland_flow)

compare_times <- rbind(sim_catchment$`0.1`$gauges_flow[11,],
                       sim_catchment$`0.5`$gauges_flow[3,],
                       sim_catchment$`1`$gauges_flow[2,])

compare_times$max <- c(max(sim_catchment$`0.1`$gauges_flow$GAUGE_01),
                       max(sim_catchment$`0.5`$gauges_flow$GAUGE_01),
                       max(sim_catchment$`1`$gauges_flow$GAUGE_01))

compare_times$timestep <- c("0.1", "0.5", "1")
print(compare_times)
```
Der Algorithmus sollte bei unterschiedlichen Zeiteinheiten, aber gleichem Simulationszeitpunkt eigentlich die selben Ergebnisse erhalten.  
Allerdings kommen bei kleineren Zeiteinheiten wohl auch mehr Regeneinheiten dazu (größere maximalwerte bei kleinerer Zeiteinheit).  
  
***  
  
#### function hy_analysis

```{r, eval = FALSE}

hy_basic_analysis <- function(DEM_path, workdir, sink_min_slope = "0.1"){
  
  library(gdalUtils)
  library(rgdal)
  library(raster)

  
  # Create output- and workfolder
  if(!file.exists(file.path(workdir))){
    dir.create(file.path(workdir), recursive = TRUE)
  }
  setwd(workdir)
  
  # if the function already were executet, just read the output RDS
  if(file.exists(paste0(workdir, "hy_results.rds"))){
    
    res <- readRDS(paste0(workdir, "hy_results.rds"))
    return(res)
    
  # if there is no hy_results.rds in the work directory, execute SAGA Moduls  
  }else{
    
    # 1. convert tif to SAGA Format
    gdalwarp(srcfile = DEM_path, dstfile = paste0(workdir, "DEM.sdat"), overwrite = TRUE, of = 'SAGA')
    
    # 2. Preprocessing the DEM: Fill Sinks (Wang & Liu)
    system(paste0("saga_cmd ta_preprocessor 4 ",
                  "-ELEV=DEM.sdat ",
                  "-FILLED=DEM_no_sinks.sdat ",
                  "-FDIR=flow_direction.sdat ",
                  "-WSHED=watershed.sdat ",
                  "-MINSLOPE=",sink_min_slope))
    
    
    # 3. Catchment areas
    system(paste0("saga_cmd garden_learn_to_program 7 ",
                  " -ELEVATION=DEM_no_sinks.sdat ",
                  " -AREA=catchment_area.sdat ",
                  " -METHOD=0"))
    
    
    # Last: convert output to tif
    # task: files besser in liste ordnen; liste benennen
    output <- list.files(workdir, pattern = "*.sgrd", full.names = TRUE)
    
    ret <- lapply(seq(length(output)),function(i){
      
      gdalwarp(srcfile = sub(".sgrd",".sdat",output[i]),
               dstfile = sub(".sgrd",".tif",output[i]),
               overwrite = TRUE, of = 'GTiff')
      
      raster(sub(".sgrd",".tif",output[i]))
      
    })
    names(ret) <- substr(list.files(workdir, pattern = "*.tif"), start = 1, stop = nchar(list.files(workdir, pattern = "*.tif"))-4)
    saveRDS(ret, file = paste0(workdir, "hy_results.rds"))
    return(ret)
    
  }
}  
```

***  

#### function overland_flow
```{r, eval = FALSE}
hy_overland_flow <- function(DEM_path,
                             gauges_path,
                             workdir,
                             time = "24",
                             steps = "0.1",
                             roughness = "0.03",
                             maxiter = "100",
                             epsilon = "0.0001",
                             precip = "0",
                             threshold = "0.0000"){
  library(raster)
  library(gdalUtils)
  
  if(!file.exists(file.path(workdir))){
    dir.create(file.path(workdir), recursive = TRUE)
  }
  setwd(workdir)
    
  system(paste0("saga_cmd sim_hydrology 1 ",
                "-DEM=", DEM_path," ",
                "-FLOW=overland_flow.sdat ",
                "-GAUGES=", gauges_path, " ",
                "-GAUGES_FLOW=gauges_flow.csv ",
                "-TIME_SPAN=", time," ",
                "-TIME_STEP=", steps, " ",
                "-ROUGHNESS=", roughness," ",
                "-NEWTON_MAXITER=", maxiter," ",
                "-NEWTON_EPSILON=",epsilon ," ",
                "-PRECIP=",precip ," ",
                "-THRESHOLD=",threshold))

  gdalwarp(srcfile = paste0(workdir, "overland_flow.sdat"),
           dstfile = paste0(workdir, "overland_flow.tif"),
           overwrite = TRUE,
           of = "GTiff")
  
  res <- list(overland_flow = raster(paste0(workdir, "overland_flow.tif")),
              gauges_flow = read.csv(paste0(workdir, "gauges_flow.csv")))
  saveRDS(res, file = paste0(workdir, "overland_flow.rds"))
  return(res)
  
}
```