---
title: "gi-ws-05-3"
author: "Charlotte Gohr, Marvin Ludwig, Klaas Rüggeberg"
date: "4 Dezember 2016"
output: html_document
---

## gi-ws-05-3 plains vs. plateaus

#### workflow:

1. DEM einlesen
2. Parameter definieren
3. DEM filtern (mean) 
4. Slope, Aspect, Curvature von beidem berechnen
5. Fuzzy Landform Element Classification für beide durchführen
6. Ungefilterte Klassifikation modal filtern
7. Als Plain oder Plateau reklassifizieren je nach Höhe im DEM
    + Ungefiltert
    + Mean Filter
    + Modal Filter

Strukturierung: Alle Konvertierungen und SAGA Module mittels einer Funktion erledigen  
So können übersichtlich mehrere DEM-tifs über eine Schleife ausgewertet werden.

Funktion muss leisten:  
* Einlesen eines DEM-tifs  
* Konvertierung mittels gdal in SAGA Format  
* Ausführen der SAGA Module  
* Durch einen Präfix überschreiben verhindern und so mehrere Dateien abarbeiten  
* Erzeugen eines Output-tifs mittels gdal  
* den Output in der aktuellen R-Session verfügbar machen  

```{r, eval = FALSE}
# gi-ws-05-3 classify plains and plateaus with a real structure using a function

library(gdalUtils)
library(rgdal)
library(raster)

# Initialise script; set filepaths and set system variables for SAGA and GDAL
source("D:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/Init.R")
path <- Init("gi", "05")

# get a list of the available functions
list.files(pattern = "[.]R$", path = path$fun, full.names = FALSE)
# source script for classification
source(paste0(path$fun, "fuzzy_land_class.R"))

# create filtered DEM with focal filter
DEM <- raster(paste0(path$gi$input, "geonode_las_dtm_01m.tif"))
DEM_filter <- focal(DEM, w = matrix(1,3,3), fun = mean)
writeRaster(DEM_filter, filename = paste0(path$gi$input, "geonode_las_dtm_01m_filter.tif"), overwrite = TRUE)


# # # # Set parameters for classification # # # #
DEM_paths <- c(paste0(path$gi$input, "geonode_las_dtm_01m.tif"),
               paste0(path$gi$input, "geonode_las_dtm_01m_filter.tif"))
# set prefixes for different DEMs to prevent overwriting files in the function
prefix <- c("norm_", "filt_")
# set patameters for 'Fuzzy Landform Element Classification'
slope_min <- "3"
slope_max <- "10"
curve_min <- "0.00000001"
curve_max <- "0.0001"


# # # # Use the sourced function to classify the normal and filtered DEM # # # #
#
# the function runs the SAGA module 'Slope, Aspect, Curvature' to create the input rasters
# for the module 'Fuzzy Landform Element Classification'
# output is the raster 'landform' from 'Fuzzy Landform Element Classificaton'
# plains have a value of 100 in the output raster

landforms <- lapply(seq(2), function(i){
  
  classify(DEM = DEM_paths[i],
           work_dir = path$gi$run,
           output_dir = path$gi$output,
           pre = prefix[i],
           T_SLOPE_MIN = slope_min,
           T_SLOPE_MAX = slope_max,
           T_CURVE_MIN = curve_min,
           T_CURVE_MAX = curve_max)
})

# save classified landform in rasters
landform_norm <- landforms[[1]]
landform_filt <- landforms[[2]]

# modal filter with the unfiltered classification
landform_mod_filt <- focal(landform_norm, w = matrix(1,7,7), fun = modal)
writeRaster(landform_mod_filt, filename = paste0(path$gi$output, "landform_mod_filt.tif"))


# Reclassify as plains or plateaus based on height of the DEM
reclass <- lapply(c(landform_norm,
                    landform_filt,
                    landform_mod_filt), function(landform){
  landform <- reclassify(landform, c(0,99,0, 99,100,1,100,200,0 ))                    
  landform[landform == 1 & DEM < 250] <- 1
  landform[landform == 1 & DEM >= 250] <- 2
  return(landform)
})

# Show results
plot(reclass[[1]], sub = "yellow = plain; green = plateau")
title("P&P unfiltered")

plot(reclass[[2]], sub = "yellow = plain; green = plateau")
title("P&P DEM filter")

plot(reclass[[3]], sub = "yellow = plain; green = plateau")
title("P&P modal filter")
```

## Show results
```{r}

# Initialise script again for the RMD
library(raster)

source("D:/university/msc-phygeo-class-of-2016-Ludwigm6/fun/Init.R")
path <- Init("gi", "05")

# load data again for the RMD
DEM <- raster(paste0(path$gi$input, "geonode_las_dtm_01m.tif"))
landform_norm <- raster(paste0(path$gi$output, "norm_class_landform.tif"))
landform_filt <- raster(paste0(path$gi$output, "filt_class_landform.tif"))
landform_mod_filt <- raster(paste0(path$gi$output, "landform_mod_filt.tif"))

# Reclassify as plains or plateaus based on height of the DEM
reclass <- lapply(c(landform_norm,
                    landform_filt,
                    landform_mod_filt), function(landform){
  # reclass as plains (value 100) and no-plains (all other values)                    
  landform <- reclassify(landform, c(0,99,0, 99,100,1,100,200,0 ))  
  # reclass as plain or plateau
  landform[landform == 1 & DEM < 250] <- 1
  landform[landform == 1 & DEM >= 250] <- 2
  return(landform)
})

# Show results
plot(reclass[[1]], sub = "yellow = plain; green = plateau")
title("P&P unfiltered")
# evaluation: many artifacts and fragmented areas 

plot(reclass[[2]], sub = "yellow = plain; green = plateau")
title("P&P DEM filter")
# evaluation: many artifacts, visible lines (maybe height lines? slope parameter too low?)

plot(reclass[[3]], sub = "yellow = plain; green = plateau")
title("P&P modal filter")
# evaluation: nearly no artifacts but unrealistic plateau sizes
# best would be a combination of both filters
```

## Function 'classify'
```{r, eval = FALSE}
# # # # SAGA Fuzzy Landform Element Classification # # # #
#
# runs the SAGA modules 'Slope, Aspect, Curavture' and 'Fuzzy Landform Element Classification'
# to create the landform classification out of a DEM
#
# # # Inputs # # #
# DEM: filepath to the elevation model [.tif]
# work_dir: directory for the 'temporary' rasters like slope etc [.sdat]
# output_dir: directory for the output raster landform [.tif]
# pre: prefix for all created data to prevent overwriting of existing files and repeatability
# T_SLOPE_MIN, T_SLOPE_MAX, T_CURVE_MIN, T_CURVE_MAX: parameters for FLEC as Text
#
# # # Output # # #
# landform raster as [.tif] in the output directory and as a raster layer in R
#

classify <- function(DEM,
                     work_dir,
                     output_dir,
                     pre,
                     T_SLOPE_MIN,
                     T_SLOPE_MAX,
                     T_CURVE_MIN,
                     T_CURVE_MAX){
  library(gdalUtils)
  library(raster)
  
  gdalwarp(srcfile = DEM, dstfile = paste0(work_dir, pre, "DEM.sdat"), overwrite = TRUE, of = 'SAGA')
  
  setwd(work_dir)
  system(paste0("saga_cmd ta_morphometry 0 ",
                "-ELEVATION=",pre,"DEM.sdat ",
                "-SLOPE=",pre,"slope ",
                "-ASPECT=",pre,"aspect ",
                "-C_PROF=",pre,"profil ",
                "-C_TANG=",pre,"tangential ",
                "-C_MINI=",pre,"minimum ",
                "-C_MAXI=",pre,"maximum ",
                "-METHOD=6 -UNIT_SLOPE=1 -UNIT_ASPECT=1"))
  
  
  system(paste0("saga_cmd ta_morphometry 25 ",
                "-SLOPE ",pre,"slope.sgrd ",
                "-MINCURV ",pre,"minimum.sgrd ",
                "-MAXCURV ",pre,"maximum.sgrd ",
                "-PCURV ",pre,"profil.sgrd ",
                "-TCURV ",pre,"tangential.sgrd ",
                "-FORM ",pre,"class_landform ",
                "-MEM ",pre,"class_membership ",
                "-ENTROPY ",pre,"class_entropy ",
                "-CI ",pre,"class_confusion ",
                "-SLOPETODEG 0 ",
                "-T_SLOPE_MIN=",T_SLOPE_MIN, " ",
                "-T_SLOPE_MAX=",T_SLOPE_MAX, " ",
                "-T_CURVE_MIN=",T_CURVE_MIN, " ",
                "-T_CURVE_MAX=",T_CURVE_MAX))
  
  
  gdalwarp(srcfile = paste0(work_dir, pre,"class_landform.sdat"), 
           dstfile = paste0(output_dir, pre,"class_landform.tif"), 
           overwrite = TRUE)
  landform <- raster(paste0(output_dir, pre,"class_landform.tif"))
  return(landform)
}

```

